---
title: "HW9"
author: "Jillian Egland"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Team members: Jillian Egland 
https://github.com/jcegland/practice4 



## Reading in data 
```{r} 
library(tidyverse)
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)


```



```{r} 
av0 = av %>% filter(Name.Alias !="") 

av1 = av0 %>% gather(key=time, value=death, c(11,13,15,17,19)) %>%
  select(Name.Alias, time, death) %>% 
  mutate(time=parse_number(time)) 
head(av1)

av2 = av0 %>% gather(key=time, value=return, c(12,14,16,18,20)) %>% 
  select(Name.Alias, time, return) %>% 
  mutate(time=parse_number(time)) 
head(av2)

av.neat=left_join(av1, av2, by = c("Name.Alias", "time"))
head(av.neat)

```


```{r} 
av0 = av %>% filter(Name.Alias !="Vance Astrovik") 

av1 = av0 %>% gather(key=time, value=death, c(11,13,15,17,19)) %>%
  select(Name.Alias, time, death) %>% 
  mutate(time=parse_number(time)) 
head(av1)

av2 = av0 %>% gather(key=time, value=return, c(12,14,16,18,20)) %>% 
  select(Name.Alias, time, return) %>% 
  mutate(time=parse_number(time)) 
head(av2)

av.neat=left_join(av1, av2, by = c("Name.Alias", "time"))
head(av.neat)



```




## Part two: Validation

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

Upload your answers and the code to the repository. Discuss and refine answers as a team.


# Validation - Jillian 
There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.  

To find the chance of a returning from death the first time, divide 44 fake deaths/total deaths 63 = 70% chance which is close to 2-in-3 chance. 
To find the chance of returning from death the second time, divide 8 fake deaths/total deaths 15 = 53% chance which is close to 50% chance  

so the statments were correct  

```{r} 
av.neat %>% group_by(time) %>% summarise(total.num = n(),
                                         real.death = sum(death == "YES"), 
                                         fake.death = sum(death == "NO"), 
                                         missing = sum(death == ""))

# number who died first time 
av.neat %>% filter(time == 1) %>% 
  summarise(total.num = n(),
            real.death = sum(return == "NO"), 
            fake.death = sum(return == "YES"), 
            total.deaths = real.death+fake.death,
            missing = sum(return == ""))

# number who died second time 
av.neat %>% filter(time == 2) %>% 
  summarise(total.num = n(),
            real.death = sum(return == "NO"), 
            fake.death = sum(return == "YES"), 
            total.deaths = real.death+fake.death,
            missing = sum(return == ""))

``` 









